---
title: "simpleRegress"
author: "Johana Quintanilla"
date: "Fall 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
simp1 <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(simp1)
summary.lm(simp1)
```

The predicted overall weight of a bear based on chest, has an adjusted R-squared value of 0.9262, and residual standard value of 33.08.

```{r}
simp2 <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(simp2)
summary.lm(simp2)
```

The predicted overall weight of a bear based on length, has an adjusted R-squared value of 0.7423, and residual standard value of 61.83.

```{r}
simp3 <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(simp3)
summary.lm(simp3)
```

The predicted overall weight of a bear based on head width, has an adjusted R-squared value of 0.6064, and residual standard value of 76.41.

```{r}
simp4 <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(simp4)
summary.lm(simp4)
```

The predicted overall weight of a bear based on head length, has an adjusted R-squared value of 0.69, and residual standard value of 67.81.

```{r}
simp5 <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(simp5)
summary.lm(simp5)
```

The predicted overall weight of a bear based on age, has an adjusted R-squared value of 0.5526, and residual standard error value of 81.47.

##   CONCLUSION AND RANKING

According to the data we collected for the adjusted R-squared and residual standard error values when we compared weight to all other numerical values, the ranking is as such:

1) WEIGHT~CHEST: RSE: 33.08 & ARS: 0.9262
2) WEIGHT~NECK: RSE: 43.91 & ARS: 0.87
3) WEIGHT~LENGTH: RSE: 61.83 & ARS: 0.7423
4) WEIGHT~HEAD LENGTH: RSE: 67.81 & ARS: 0.69
5) WEIGHT~HEAD WIDTH: RSE: 76.41 & ARS: 0.6064
6) WEIGHT~AGE: RSE: 81.47 & ARS: 0.5526
